distr shiftate + troncate

### 2016-09-14 docs/article-shift+trunc.Rmd
articolo distr trunc + shift + prove

* probability distribution functions p<dist>(q) with q vector of quantiles;
* density functions d<dist>(x) with x vector of quantiles;
* quantile functions q<dist>(p) with p vector of probabilities;
* random number generation r<dist>(n) with n number of observations.

where <dist> indicates the distribution family.

@@

```{r}
x <- seq(6, 14, len = 100)
std_density <- dnorm(x = x, mean = 10, sd = 1)
trunc_density <- dtnorm(x = x, mean = 10, sd = 1, L = 9, U = 11)

@@

@@

# Distribution function for truncated and shifted or for shifted and truncated distributions

@@

We have to define also the others functions for truncated distribution, as they are necessary for the further step:

@@

We have to define also the others functions for shifted distribution, as they are necessary for the further step:

@@

So `ptsnorm()` is the function that computes the density function of `stnorm`, which is the truncated shifted normal distribution.

@@

We can conclude that the previous functiona are completely different.

>>>>

* probability distribution functions p< dist >(q) with q vector of quantiles;
* density functions d< dist >(x) with x vector of quantiles;
* quantile functions q< dist >(p) with p vector of probabilities;
* random number generation r< dist >(n) with n number of observations.

where < dist > indicates the distribution family.

@@

```{r, message=FALSE}
x <- seq(6, 14, len = 100)
std_density <- dnorm(x = x, mean = 10, sd = 1)
trunc_density <- dtnorm(x = x, mean = 10, sd = 1, L = 9, U = 11)

@@

```{r}
args(dnorm)
args(dsnorm)
```

@@

# Distribution functions for truncated and shifted or for shifted and truncated distributions

@@

We have to define also the others functions for truncated distribution, as they are necessary implicitly for the further step:

@@

We have to define also the others functions for shifted distribution, as they are implicitly necessary for the further step:

@@

So `dtsnorm()` is the function that computes the density function of `stnorm`, which is the truncated shifted normal distribution.

@@

We can conclude that the previous functions are completely different.

Let us analyze the other distributions functions:

* probability distribution function of truncated-shifted normal distribution and of shifted-truncated normal distribution:

```{r}
pstnorm <- pshift("tnorm")
ptsnorm <- ptruncate("snorm")
std_distr <- pnorm(q = x, mean = 10, sd = 1)
st_distr <- pstnorm(q = x, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
ts_distr <- ptsnorm(q = x, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
```

```{r}
df <- data.frame(x=x, 
                 type = factor(c(rep("std_distr", times=length(x)),
                    rep("st_distr", times=length(x)),
                    rep("ts_distr", times=length(x))),
                    levels = c("std_distr", "st_distr", "ts_distr")), 
                  value = c(std_distr, st_distr, ts_distr))

pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
  geom_line(size=1) +
  scale_color_manual(values = c("std_distr" = "cyan3", 
                                "st_distr" = "#89FF99", 
                                "ts_distr" = "#FF89B4")) +
  xlab("Value") + ylab("Density") +
  ggtitle("Distribution plot for transformed and not-transformed Normal distributions") +
  theme(legend.title=element_blank(), legend.text= element_text(size=10),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

* quantile function of truncated-shifted normal distribution and of shifted-truncated normal distribution:

```{r}
qstnorm <- qshift("tnorm")
qtsnorm <- qtruncate("snorm")
x <- 1:999/1000
std_quant <- qnorm(p = x, mean = 10, sd = 1)
st_quant <- qstnorm(p = x, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
ts_quant <- qtsnorm(p = x, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
```

```{r}
df <- data.frame(x=x, 
                 type = factor(c(rep("std_quant", times=length(x)),
                    rep("st_quant", times=length(x)),
                    rep("ts_quant", times=length(x))),
                    levels = c("std_quant", "st_quant", "ts_quant")), 
                  value = c(std_quant, st_quant, ts_quant))

pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
  geom_line(size=1) +
  scale_color_manual(values = c("std_quant" = "cyan3", 
                                "st_quant" = "#89FF99", 
                                "ts_quant" = "#FF89B4")) +
  xlab("Value") + ylab("Density") +
  ggtitle("Distribution plot for transformed and not-transformed Normal distributions") +
  theme(legend.title=element_blank(), legend.text= element_text(size=10),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

* random generation function of truncated-shifted normal distribution and of shifted-truncated normal distribution:

```{r}
rstnorm <- rshift("tnorm")
rtsnorm <- rtruncate("snorm")
std_rand <- rnorm(n=1000, mean = 10, sd = 1)
st_rand <- rstnorm(n=1000, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
ts_rand <- rtsnorm(n=1000, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
```

```{r}
df <- data.frame(x=1:1000, 
                 type = factor(c(rep("std_rand", times=1000),
                    rep("st_rand", times=1000),
                    rep("ts_rand", times=1000)),
                    levels = c("std_rand", "st_rand", "ts_rand")), 
                  value = c(std_rand, st_rand, ts_rand))

pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
  geom_line(size=1) +
  scale_color_manual(values = c("std_rand" = "cyan3", 
                                "st_rand" = "#89FF99", 
                                "ts_rand" = "#FF89B4")) +
  xlab("Value") + ylab("Density") +
  ggtitle("Distribution plot for transformed and not-transformed Normal distributions") +
  theme(legend.title=element_blank(), legend.text= element_text(size=10),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

<!--
pensare ad un modo diverso per spiegare la cosa perchè da questi grafici nn si capisce niente
-->
## Other Examples

Let us see how it the other distribution functions works with other distributions.

Let us consider Poisson distribution and distribution function.

First of all functions for truncated distribution and for shifted distributio have to be defined:

```{r}
dtpois <- dtruncate("pois")
ptpois <- ptruncate("pois")
qtpois <- qtruncate("pois")
rtpois <- rtruncate("pois")

dspois <- dshift("pois")
pspois <- pshift("pois")
qspois <- qshift("pois")
rspois <- rshift("pois")
```
